Low Infant Birth Weight in Brazil

### 2024-06-24 products/manuscript/Manuscript_Final.qmd
Revisions
Updated manuscript and Assets Readme and Master Readme

---
title: "Low Infant Birth Weight in Brazil"
subtitle: "A Historical Data Analysis Approach"
author: Cassia Roth
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
    bibliography: ../../assets/Mada_Bib.bib
    csl: ../../assets/vancouver.csl
---

@@

# Summary/Abstract

This cross-sectional study uses linear regression models to understand disparities in infant birth weight in relation to maternal skin color (ancestry), age, gestational status (parity), and nationality in Rio de Janeiro, Brazil's first public maternity hospital, the Maternidade Laranjeiras (now Maternidade Escola) in the 1920s. I hypothesize that infants born to women of color (either Black or "*preta*" or mixed-race or "*parda*") will have lower birth weights than infants born to White women, whether Brazilian or immigrant, given the historical legacies of slavery, only abolished in 1888, on population health in Brazil.

@@

>>>>

---
title: "Low Infant Birth Weight in Brazil"
subtitle: "A Historical Data Analysis Approach"
author: Cassia Roth
date: "`r Sys.Date()`"
format:
  docx:
    toc: false
    number-sections: true
    highlight-style: github
    bibliography: ../../assets/Mada_Bib.bib
    csl: ../../assets/chicago-fullnote-bibliography.csl
---

@@

# Abstract

This cross-sectional study uses linear regression models to understand disparities in infant birth weight in relation to maternal skin color (ancestry), age, gestational status (parity), and nationality in Rio de Janeiro, Brazil's first public maternity hospital, the Maternidade Laranjeiras (now Maternidade Escola) in the 1920s. Infants born to women of color (either Black or "*preta*" or mixed-race or "*parda*") had lower birth weights than infants born to White women, whether Brazilian or immigrant. I hypothesize these differences were related to the historical legacies of slavery, only abolished in 1888, on population health in Brazil.

@@

At 4 in the morning on November 17, 1922, Virginia Maria da Conceição, a twenty-six-year-old Black ("*preta*") woman gave birth to a stillborn daughter in Rio de Janeiro, Brazil's first public maternity hospital, the Maternidade Laranjeiras (now called Maternidade Escola). This was Conceição's fifth birth; prior, she had delivered three full-term infants vaginally (described as "naturally" in the clinical notes) and one premature infant, born at seven months gestational age. According to the notes, Conceição had had a "good" pregnancy. After arriving at the hospital, she remained in the first stage of labor ("dilatação") for five hours. The second stage of labor ("expulsão") lasted only thirty minutes. The placenta was delivered ten minutes later. Despite the apparently easy pregancy and delivery, Conceição would not take home a new baby; rather, "hereditary syphilis," as the fetal autopsy noted, was the cause of the stillbirth. The unnamed infant weighed 3,180 grams at birth, a normal weight for a full-term infant. [PAGE 542-4, Dezembro 1922, 16 no. 12]

Virginia Maria da Conceição's experience with stillbirth is one of the 2845 recorded clinical visits to Maternidade Laranjeiras between June 1922 and May 1926 that form the source base for this article. It is also the only one to include the patient's name. For historians of medicine and public health, this breach of privacy is a boon to social historical methods, which analyze these histories from the "patient's point of view" CITE. For contemporary medical practitioners, the inclusion of Conceição's name may come as a shock, as patient privacy is a cornerstone of medical ethics. I use her story here as a jumping off point to explore the relationship between social determinants of health and infant birthweight in a former slave society.
